### 2019-01-24 Thesis_Latex/Thesis/Chapter4/Chapter4.tex

\chapter{Moving Obstacle Avoidance}
\section{Problem Formulation}
\section{Design of Adaptive MPC}
\section{Simulation Results}