### 2019-01-24 Thesis_Latex/Thesis/Chapter5/Chapter5.tex

\chapter{Lane Following}
\section{Problem Formulation}
\section{Design of Adaptive MPC}
\section{Simulation Results}