### 2019-01-24 Thesis_Latex/Thesis/Conclusions/Conclusions.tex

\chapter*{Conclusions and Future Work}
\addcontentsline{toc}{chapter}{Conclusions and Future Work}