### 2019-02-04 Thesis_Latex/Paper/figure/flowchart/flowchart.tex

\documentclass[preview,float]{standalone}

\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\begin{document}
\pagestyle{empty}

% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=green!20, 
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
    minimum height=2em]

\begin{tikzpicture}[node distance = 3.3cm, auto]
    % Place nodes
    \node [decision] (detection) {Obstacle 
    detected?}; 
    \node [cloud,right of=detection] (start) {Start};
    \node [decision,below of=detection] (multiObstacles) {Multiple Obstacles detected?};
	\node [block,left of=detection] (reference) {Follow Path: stick to the lane};
    \node [block,below of=multiObstacles] (distance) {Distance to closest obstacle};
    \node [block,right of=multiObstacles] (distances) {Distance between the obstacles};
    \node [decision,below of=distances](space) {Free Space for Overtaking?};
    \node [block,right of=space](braking){Braking/ Ostacle Velocity};
    \node [block,below of=space](free){Overtaking on free lane};
    \node [decision,below of=distance] (decisionOvertaking) {Left w.r.t vehicle?};
    \node [block, below of=decisionOvertaking] (rightOver) {Right Overtaking};     
    \node [block, left of=decisionOvertaking] (leftOver) {Left Overtaking};

  %  \node [block] (init) {initialize model};
   % \node [cloud, left of=init] (expert) {expert};
   % \node [cloud, right of=init] (system) {system};
   % \node [block, below of=init] (identify) {identify candidate models};
%    \node [block, below of=identify] (evaluate) {evaluate candidate models};
 %   \node [block, left of=evaluate, node distance=3cm] (update) {update model};
  %  \node [decision, below of=evaluate] (decide) {is best candidate better?};
   % \node [block, below of=decide, node distance=3cm] (stop) {stop};
    % Draw edges
    \path [line] (start) -- (detection);
    \path [line] (detection) -- node {yes} (multiObstacles);
    \path [line] (detection) -- node {no} (reference);
    \path [line] (multiObstacles) -- node {no} (distance);
    \path [line] (multiObstacles) -- node {yes} (distances);
    \path [line] (distances)--(space);
    \path [line] (space) -- node{no}(braking);
    \path [line] (space) -- node{yes} (free);
    \path [line] (braking) |- (distances);
    \path [line] (distance) -- (decisionOvertaking);
    \path[line] (decisionOvertaking) -- node {no}(leftOver);
    \path[line] (decisionOvertaking) -- node{yes} (rightOver);
\end{tikzpicture}

\end{document}