### 2019-03-06 Thesis_Latex/Thesis/Acknowledgements/Acknowledgements.tex
update thesis

\chapter*{Ringraziamenti/Acknowledgements}
\thispagestyle{empty}
dedica a parenti e amici

The research work for this thesis was carried out at the Laboratory for Automation and Robotics (LAR) in the Department of Mechanical Engineering of Aveiro (Portugal), during a period of 5 months as an exchange student.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.50\textwidth]{../figure/logo_ua.png}
	\label{fig:logo_ua}
\end{figure}

I would like to express my gratitude to Professor Vitor Santos, for welcoming me at the Laboratory for Automation and Robotics and making possible my experience of study at the University of Aveiro. His help has been invaluable both personally and academically. Moreover I would like to thank Professor Angelo Cenedese for his advices and his help during this work.